Global patterns of forest autotrophic carbon fluxes

### 2020-03-01 manuscript/manuscript draft.Rmd
worked on methods

*Model specification.*

{**from Helene: It was frequently unclear to me whether the influences of other climate variables were being considered in multivariate or univariate analyses.  This needs to be clear.  I can see reasons to do it either way, but they need to be differently motivated and clearly presented.**}

The effects of climate and latitude on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. First, the individual effect of each climate variable on each flux variable was modelled by specifying the climate variable as a fixed effect in a univariate model. For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a second-order polynomial term. AIC values were calculated for the models and used to select the best model. If the best model included a polynomial term, the shape of the polynomial relationship was considered. If the shape of the relationship *made biological sense*, and was a significant improvement on the linear relationship (deltaAIC >2), we accepted the polynomial as the best model. If not, we ran the linear model as the final model. $R^2$ values were calculated for the best model. All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified. In addition, slope coefficients were calculated for the linear models. {**This paragraph needs work - polynomial models were rejected if they showed an exponential increase in productivity with climate (e.g. showed GPP increasing exponentially with latitude - needs work on how to word this**}

To test for a potential influence of altitude, models were also run with site altitude included as a second fixed effect. These multivariate models were compared against univariate models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models. Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power. As a result, it was decided not to include altitude as a fixed term in the final models.

To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area. Geographically proximate sites are clustered into geographic areas, with proximity defined using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].

Models were run for total annual flux against annual climate variables, and for fluxes standardized by growing season length, defined as annual flux/length of growing season, against growing season climate variables. For analyses on data within the growing season, only linear models were specified.

{**Note from Helene on the following paragraph: this makes it sound like analyses from H3 were about variation remaining after accounting for temperature, but analyses below don't seem to do that.**}   
To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified. To ensure that models were biologically meaningful, the terms included in the models tested built on results from the univariate models. Modelling of individual climate variables identified that the best predictors of carbon fluxes were variables related to temperature. We therefore decided to include mean annual temperature as a term in all multivariate models. We first modelled the interaction effect between mean annual temperature and mean annual precipitation, in order to capture climate variation along the axes of temperature and water availability. Models were tested for a significant interactive effect and a significant additive effect. We then explored whether any other climate variable, in combination with mean annual temperature, could significantly improve on the combination of mean annual temperature and mean annual precipitation. In specifying the range of models to test, climate variables which were strongly correlated with temperature were dropped, in order to capture the greatest range of variation in climate. For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect. As before, plot nested within geographic area was included as a random effect. Altitude was not considered. AIC values were calculated for the models, and used to compare models. Models were considered to be significantly better than the baseline MAT*MAP model if:  
  i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by >0.05

*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes. To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component. The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux. Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015]. Stacked plots were generated for:  
1. GPP = NPP + R~auto~  
2. NPP = ANPP + BNPP  
3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
4. Total belowground carbon flux = BNPP + R~root~

*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~. For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated. The ratios were regressed against latitude and climate variables, using the linear model specified above. Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions. To account for this, models were re-run using only data from sites $\le$ 1000m.

@@

In total, we analyzed 1228 records from 9 C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years. These records represented a total of 154 distinct geographic areas (Fig. 1, Table 2), across all forested biogeographic and climate zones.

>>>>

*Analyses*

The effects of latitude and climate on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018]. We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the change in AIC was generally small (often <2) relative to models excluding altitude. Hypotheses were accepted if the AIC difference ($\Delta$AIC) between a model including the fixed effect of interest and a corresponding null model excluding that fixed effect exceeded 2.0. All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects, **unless otherwise specified**. **Becky, are there any instances of this?** Specific analyses are as described below.

We first examined the relationship between latitude and FACF (Q1; Table 1). We tested models with latitude as a linear and term (corresponding null: model without latitude) and as a second-order polynomial term (corresponding null: model with latitude as a linear term). If the shape of the relationship [**made biological sense**](https://github.com/forc-db/Global_Productivity/issues/64), and was a significant improvement on the linear relationship ($\Delta$AIC >2), we accepted the polynomial as the best model. We also examined relationships among fluxes across latitude, testing whether sums of component fluxes matched the larger fluxes and whether C allocation varied with latitude, as specified below.

**work on this paragraph** 
*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes. To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component. The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux. Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015]. Stacked plots were generated for:  
1. GPP = NPP + R~auto~  
2. NPP = ANPP + BNPP  
3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
4. Total belowground carbon flux = BNPP + R~root~

**work on this paragraph**   
*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~. For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated. The ratios were regressed against latitude and climate variables, using the linear model specified above. Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions. To account for this, models were re-run using only data from sites $\le$ 1000m.

**work on this paragraph**  
**We next examined the relationships of FACF to climate variables (Q2-Q4; Table 1)...** As with latitude, we tested booth linear and polynomial fits for each climate variable.

{**Note from Helene on the following paragraph: this makes it sound like analyses from H3 were about variation remaining after accounting for temperature, but analyses below don't seem to do that.**}  
**work on this paragraph**  
To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified. To ensure that models were biologically meaningful, the terms included in the models tested built on results from the univariate models. Modelling of individual climate variables identified that the best predictors of carbon fluxes were variables related to temperature. We therefore decided to include mean annual temperature as a term in all multivariate models. We first modelled the interaction effect between mean annual temperature and mean annual precipitation, in order to capture climate variation along the axes of temperature and water availability. Models were tested for a significant interactive effect and a significant additive effect. We then explored whether any other climate variable, in combination with mean annual temperature, could significantly improve on the combination of mean annual temperature and mean annual precipitation. In specifying the range of models to test, climate variables which were strongly correlated with temperature were dropped, in order to capture the greatest range of variation in climate. For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect. As before, plot nested within geographic area was included as a random effect. Altitude was not considered. AIC values were calculated for the models, and used to compare models. Models were considered to be significantly better than the baseline MAT*MAP model if:  
  i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by >0.05

**work on this paragraph**  
Models were run for total annual flux against annual climate variables, and for fluxes standardized by growing season length, defined as annual flux/length of growing season, against growing season climate variables. For analyses on data within the growing season, only linear models were specified.

@@

In total, we analyzed 1228 records from nine FACF variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years. These records represented a total of 154 distinct geographic areas (Fig. 1, Table 2), across all forested biogeographic and climate zones.
